Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case

### 2019-03-10 ICML-submission/rebuttal.txt

We thank all reviewers for their insightful comments. We will make a pass,
and incorporate all the detailed comments in the final version of the paper.

We believe that extending the algorithms to handle the general
contextual bandit learning setting (where each example is associated
with a cost vector) requires nontrivial modifications to the algorithm;
a first step towards this question would be to establish
regret guarantees for when the examples shown is still multiclass, but
linearly nonseparable - we leave this for future work.

Regarding Appendix G, we agree that it is not related to the main story
 of the paper. Nevertheless, we think that this is a baseline that has not
 appeared in the literature and we include it for completeness.

We will add some discussions on achievable mistake bounds in the weakly
linearly separable case,
and its interaction with computational efficiency requirements.
It is known that in (Daniely and Herbertal, 2013)
that when 1/gamma^2 < d, any algorithm must make \Omega(K / \gamma^2) mistakes,
and when 1/gamma^2 > K^3 d, any algorithm must make \Omega( d K^2 ) mistakes.
In addition, (Daniely and Herbertal, 2013) gives a computationally inefficient
 algorithm that has a mistake upper bound of \tilde{O}( \min( K/\gamma^2, d K^2) );
 the algorithm involves a key step of computing (margin-based) Littlestone dimensions of subsets of multiclass linear classifiers, which is not known how to do efficiently.

### 2019-03-10 ICML-submission/rebuttal.txt

We thank all reviewers for their insightful comments. We will make a pass, and incorporate all the detailed comments in the final version of the paper.

-----
Reviewer #2
We believe that extending the algorithms to handle the general contextual bandit learning setting (where each example is associated with a cost vector) requires nontrivial modifications to the algorithm; a first step towards this question would be to establish regret guarantees for when the examples shown is still multiclass, but linearly nonseparable - we leave this for future work.

@@

Regarding Appendix G, we agree that it is not related to the main story of the paper. Nevertheless, we think that this is a baseline that has not appeared in the literature and we include it for completeness.

In Appendix F, the Banditron's boundary is plot with the version with the largest exploration rate (the 0.02 one in Figure 4 and 5). This version explores the most, and gives the most accurate decision boundary. We would add more explanation on this in the final version.

-----
Reviewer #3 and #5
We will add some discussions on achievable mistake bounds in the weakly linearly separable case, and its interaction with computational efficiency requirements. It is known that in (Daniely and Herbertal, 2013) that when 1/gamma^2 < d, any algorithm must make \Omega(K / \gamma^2) mistakes, and when 1/gamma^2 > K^3 d, any algorithm must make \Omega( d K^2 ) mistakes. In addition, (Daniely and Herbertal, 2013) gives a computationally inefficient algorithm that has a mistake upper bound of \tilde{O}( \min( K/\gamma^2, d K^2) ); the algorithm involves a key step of computing (margin-based) Littlestone dimensions of subsets of multiclass linear classifiers, which is not known how to do efficiently.

-----
Reviewer #5
i. We leave the non-separable case as an important future direction. Using our current analysis framework, we believe that there will still involve the quasi-polynomial factors on K and \gamma; and the bound will only depend on L, the number of mistakes made by the best linear classifier, instead of T. To derive this, one can first re-derive our Theorem 2 for the non-separable case, and then again use the kernel trick.

ii. "Efficient" means that run-time is polynomial in K, d, \gamma, T. This follows the typical definition in online learning literature, but we will more formally define it in the final version. Since J_i^{(t)} only includes the samples on which the i-th binary perceptron makes mistakes, in the separable case, their number will be finite based on perceptron's guarantee. Therefore, it can be much better than O(t) in the separable case.

iii. Our work indeed only improves Kakade et al. and Daniely & Helbert et al.'s algorithms in efficiency, but not in mistake bound. Why Daniely and Herbertal's algorithm is inefficient is answered above.

iv. Larger-scale experiment on real data is also a future work. Currently our experiment section is more for verifying that the algorithm indeed works as expected. We will improve our figure based on your suggestion.

v. See above for the answer.

### 2019-03-12 ICML-submission/rebuttal.txt

-----
Reviewer #2
We believe that extending the algorithms to handle 
the general contextual bandit learning setting (where 
each example is associated with a cost vector) requires 
nontrivial modifications to the algorithm; a first step 
towards this question would be to establish regret 
guarantees for when the examples shown is still multiclass, 
but linearly nonseparable - we leave this for future work.

Regarding Appendix G, we agree that it is not related 
to the main story of the paper. Nevertheless, we think 
that this is a baseline that has not appeared in the 
literature and we include it for completeness.

In Appendix F, the Banditron's boundary is plot with the 
version with the largest exploration rate (the 0.02 one 
in Figure 4 and 5). This version explores the most, and 
gives the most accurate decision boundary. We would add 
more explanation on this in the final version.

-----
Reviewer #4
Thanks for the clarification suggestions. We will incorporate 
them into our final version.

@@

-----
Reviewer #5
i. We leave the non-separable case as an important future 
direction. Using our current analysis framework, we believe 
that there will still involve the quasi-polynomial factors 
on K and \gamma; and the bound will only depend on L, the 
number of mistakes made by the best linear classifier, instead 
of T. To derive this, one can first re-derive our Theorem 2 
for the non-separable case, and then again use the kernel trick.

ii. "Efficient" means that run-time is polynomial in K, d, 
\gamma, T. This follows the typical definition in online 
learning literature, but we will more formally define it in 
the final version. Since J_i^{(t)} only includes the samples 
on which the i-th binary perceptron makes mistakes, in the 
separable case, their number will be finite based on perceptron's 
guarantee. Therefore, it can be much better than O(t) in the 
separable case.

iii. Our work indeed only improves Kakade et al. and 
Daniely & Helbert et al.'s algorithms in efficiency, but not 
in mistake bound. Why Daniely and Herbertal's algorithm is 
inefficient is answered above.

iv. Indeed, currently our experiment section is more for 
verifying that the algorithm is working as expected. As a future
work, we will test our algorithm on larger-scale and real problems.

### 2019-03-13 ICML-submission/rebuttal.txt

We thank all reviewers for their helpful comments.  We will incorporate all
comments in the final version of the paper.

-----
Reviewer #2

Regarding Appendix G, we agree that it is not related to the main story of the
paper. However, it is an important example of an "ignorant" algorithm and the
lower bound that we prove in Appendix I applies to it.

In Appendix F, the Banditron's boundary is plotted with the version with
exploration rate of 0.02. This version explores the most, and gives the most
accurate decision boundary. We would add more explanation on this in the final
version.

-----
Reviewer #4

(1) The problem was motivated in the Banditron paper (Kakade et al. 2008). One
application area is web content optimization and web advertising.

(2) The phenomenon of finite number of mistakes is explained by Theorem 2 and
Corollary 6.

-----
Reviewer #3

The lower bound for strongly separable case (Theorem 3) implies the same lower
bound for the weakly separable case. This lower bound is actually tight, since
there exists an (inefficient) algorithm that makes at most O(K / \gamma^2)
mistakes due to Daniely and Herbertal, 2013 (see their Theorem 2).  We are not
aware of any mistake lower bound for computationaly efficient algorithms under
the weakly separable assumption. We will add a discussion in the paper.

-----
Reviewer #5

i. There has been substanstial amount of work on the non-separable case
as discussed in Section 2 (lines 128--137). Our paper focuses
on the remaining open case when the data is linearly separable; there
was non known computationaly efficient algorithm with finite mistake bound.

ii. "Efficient" means that run-time is polynomial in K, d, 1/\gamma, T. We will
define this explicitly in the paper.  Since J_i^{(t)} only includes the samples
on which the i-th binary perceptron makes mistakes, in the separable case,
their number will be finite based on perceptron's guarantee. The expected size of
J_i^{(t)} is upper bounded by it min(t, (1/gamma')^2).

iii. Our work indeed only improves Kakade et al. and Daniely & Helbert et al.'s
algorithms in efficiency, but not in mistake bound.

Daniely and Herbertal's algorithm is inefficient because it involves a key step
of computing (margin-based) Littlestone dimensions of subsets of multiclass
linear classifiers, which is not known how to do efficiently.

iv. We will add error bars to the plots. The focus of this submission is
largely theoretical and the goal of the experiments is to verift that the
algorithms work as expected.

v. See the answer for Reviewer #3.

### 2019-03-14 ICML-submission/rebuttal.txt

We thank all reviewers for their helpful comments.  We will incorporate all
comments in the final version of the paper.

-----
Reviewer #2

Regarding Appendix G, we agree that it is not related to the main story of the
paper. However, it is an important example of an "ignorant" algorithm and the
lower bound that we prove in Appendix I applies to it.

In Appendix F, Banditron's decision boundary is plotted with the version with
exploration rate of 0.02. This version explores the most, and gives the most
accurate decision boundary. We will add more explanation on this in the final
version.

@@

(1) The problem was motivated in the Banditron paper (Kakade et al. 2008). One
application area is web content optimization and web advertising.

(2) The phenomenon of finite number of mistakes is explained by Theorem 2 and
Corollary 6.

-----
Reviewer #3

The lower bound for strongly separable case (Theorem 3) implies the same lower
bound for the weakly separable case. This lower bound is actually tight, since
there exists an (inefficient) algorithm that makes at most O(K / \gamma^2)
mistakes due to Daniely and Herbertal, 2013 (see their Theorem 2).  We are not
aware of any mistake lower bound for computationally efficient algorithms under
the weakly separable assumption. We will add a discussion in the paper.

@@

-----
Reviewer #5

i. There has been substantial amount of work on the non-separable case
as discussed in Section 2 (lines 128--137). Our paper focuses
on the remaining open case when the data is linearly separable; there
was no known computationally efficient algorithm with finite mistake bound before
our work.

ii. "Efficient" means that the run-time is polynomial in K, d, 1/\gamma, T. We will
define this explicitly in the paper.  Since J_i^{(t)} only includes the samples
on which the i-th binary perceptron makes mistakes, its size is upper bounded by t.
This implies that the algorithm runs in polynomial time.

iii. Our work indeed only improves Kakade et al. and Daniely & Helbertal's
algorithms in computational efficiency, but not in mistake bound.
 - Kakade et al.'s algorithm is computationally inefficient because it requires
enumerating over a grid of O(1/\gamma)^kd linear classifiers.
 - Daniely & Helbertal's algorithm is computationally inefficient because it involves a key step
of computing (margin-based) Littlestone dimensions of subsets of multiclass
linear classifiers, which is not known how to do efficiently.

iv. We will add error bars to the plots. The focus of this submission is
largely theoretical and the goal of the experiments is to verify that the
algorithms work as expected.

### 2019-03-14 ICML-submission/rebuttal.txt

Regarding Appendix G, we agree that it is not related to the main story of the
paper. However, it is an important example of an "ignorant" algorithm and the
lower bound that we prove in Appendix I applies to it.

In Appendix F, Banditron's decision boundary is plotted with the version with
exploration rate of 0.02. This version explores the most, and gives the most
accurate decision boundary. We will add more explanation on this in the final
version.

-----
Reviewer #4

(1) The problem was motivated in the Banditron paper (Kakade et al., 2008). One
application area is web content optimization and web advertising.

(2) The phenomenon of finite number of mistakes made by our algorithms with
different settings of kernel functions is explained by Theorem 2 and Corollary 6.

@@

-----
Reviewer #3

The lower bound for strongly separable case (Theorem 3) implies the same lower
bound for the weakly separable case. This lower bound is actually tight, since
there exists an (inefficient) algorithm that makes at most O(K / \gamma^2)
mistakes due to Daniely & Helbertal, 2013 (see their Theorem 2).  We are not
aware of any mistake lower bound for computationally efficient algorithms under
the weakly separable assumption. We will add a discussion in the paper.

@@

-----
Reviewer #5

i. There has been substantial amount of work on the non-separable case
as discussed in Section 2 (lines 128--137). Our paper focuses
on the remaining open case when the data is linearly separable; there
was no known computationally efficient algorithm with finite mistake bound before
our work.

ii. "Efficient" means that the run-time is polynomial in K, d, 1/\gamma, T. We will
define this explicitly in the paper.  Since J_i^{(t)} only includes the samples
on which the i-th binary perceptron makes mistakes, its size is upper bounded by t.
This implies that the algorithm runs in polynomial time.

iii. Our work indeed only improves Kakade et al. and Daniely & Helbertal's
algorithms in computational efficiency, but not in mistake bound.
 - Kakade et al.'s algorithm is computationally inefficient because it requires
enumerating over a grid of O(1/\gamma)^kd linear classifiers.
 - Daniely & Helbertal's algorithm is computationally inefficient because it involves a key step
of computing (margin-based) Littlestone's dimensions of subsets of multiclass
linear classifiers, which is not known how to do efficiently.

### 2019-05-03 ICML-camera-ready/appendix-experiment.tex
add camera-ready folder

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_banditron_points.pdf}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \centering
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_banditron_points.pdf}
         \caption{Weakly separable case}
    \end{subfigure}
    \captionsetup{justification=centering}
    \caption{\textsc{Banditron}'s final decision boundaries}
     \label{fig:banditron-points}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_linear_ova_points.pdf}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \centering
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_linear_ova_points.pdf}
         \caption{Weakly separable case}
    \end{subfigure}
    \captionsetup{justification=centering}
    \caption{Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}'s final decision boundaries}
    \label{fig:linearova-points}
\end{figure}
%Our Algorithm with linear kernel (
%To better understand how each algorithm works,
%Our Algorithm with rational kernel (

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_rational_ova_points.pdf}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \centering
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_rational_ova_points.pdf}
         \caption{Weakly separable case}
    \end{subfigure}
    \captionsetup{justification=centering}
    \caption{Algorithm~\ref{algorithm:kernelized} (with rational kernel)'s final decision boundaries}
    \label{fig:rationalova-points}
\end{figure}

>>>>

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_banditron_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \centering
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_banditron_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \captionsetup{justification=centering}
    \caption{\textsc{Banditron}'s final decision boundaries}
     \label{fig:banditron-points}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_linear_ova_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \centering
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_linear_ova_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \captionsetup{justification=centering}
    \caption{Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}'s final decision boundaries}
    \label{fig:linearova-points}
\end{figure}
%Our Algorithm with linear kernel (
%To better understand how each algorithm works,
%Our Algorithm with rational kernel (

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_rational_ova_points}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \begin{subfigure}[b]{0.25\textwidth}
        \captionsetup{justification=centering}
        \centering
        \includegraphics[width=\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_rational_ova_points}
         \caption{Weakly separable case}
    \end{subfigure}
    \captionsetup{justification=centering}
    \caption{Algorithm~\ref{algorithm:kernelized} (with rational kernel)'s final decision boundaries}
    \label{fig:rationalova-points}
\end{figure}
